\LARGE \bf The Battle for Filter Supremacy: A Comparative Study of the Multi-State Constraint Kalman Filter and the Sliding Window Filter

### 2015-03-03 Report/project_report.tex
Whoops, forgot to add the report

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Accurate and consistent egomotion estimation is an important part of autonomous navigation.
For this task, the combination of visual and inertial sensors is an inexpensive, compact, and complementary hardware suite that can be used on many ground and aerial vehicles.
In this work, we compare two modern approaches to egomotion estimation: the Multi-State Constraint Kalman Filter (MSCKF) and the Sliding Window Filter (SWF).
Both filters use an Inertial Measurement Unit (IMU) to estimate the motion of a vehicle and then correct this estimate with observations of salient features from a monocular camera.
While the SWF estimates feature positions as part of the filter state itself, the MSCKF optimizes feature positions in a separate procedure without including them in the filter state.
We present experimental comparisons of the MSCKF and SWF on data from a moving hand-held sensor rig, as well as several traverses from the KITTI dataset.
In particular, we compare the accuracy of the two filters and analyze the effect of feature track length and feature density on the performance of each.
In general, our results show the SWF to be at least as accurate the MSCKF, however the MSCKF is significantly more efficient and improves in accuracy as more features are tracked. 
\end{abstract}

@@

The question, then, is how best to fuse measurements from these two sensor types to arrive at an accurate estimate of a vehicle's motion over time.
This problem is often complicated by the absence of a known map of features from which the camera can generate measurements.
Any solution must therefore solve a Simultaneous Localization and Mapping (SLAM) problem, although the importance placed on the mapping component may vary from algorithm to algorithm.
In this work we compare and contrast two modern solutions to the visual-inertial SLAM problem, the Sliding Window Filter (SWF) and the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}, on data from a moving hand-held sensor rig, as well as several traverses from the KITTI dataset \cite{Geiger:2013:IJRR}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sliding Window Filter (SWF)} \label{sec:SWF}
In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and feature positions.
The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton (GN) optimization over a sliding window of $K$ poses $\Vec x_{k\in[1,K]}$ observing $M$ features $f_{j\in[1,M]}$: 
\begin{equation}
\Vec{x} := \bbm \Vec x_0^T & \hdots & \Vec x_K^T & {\Vec {p}_G^{f_1G \hspace{0.3em} T}} & \hdots & {\Vec {p}_G^{f_MG \hspace{0.3em} T}} \ebm ^T.
\end{equation}
In our notation, $\Vec p^{BA}_A$ is the vector from the origin of frame $\mathcal{F}_A$ to the origin of frame $\mathcal{F}_B$ expressed in $\mathcal{F}_A$, and $\Vec C_{BA}$ is the rotation matrix from $\mathcal{F}_A$ to $\mathcal{F}_B$.

@@

An important advantage of the SWF is that its computational cost depends on the number of features in the current window rather than the number of features in the entire map.
By varying the spatial or temporal extent of the sliding window, the computational cost of the algorithm can be tailored to fit a given compute envelope, which makes the algorithm suitable for online operation over paths of arbitrary length.

However, the hard cut-off of the SWF may result in only some measurements of a particular feature contributing to the optimization.
As a result, the filter may not maximally constrain some vehicle poses, and hence localization may be less accurate than we could expect from the full batch solution.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Multi-State Constraint Kalman Filter (MSCKF)} \label{sec:MSCKF}
The Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
The key idea of the MSCKF is to maintain a variable window of vehicle poses and to simultaneously update each pose in the window using batch-optimized estimates of features that are visible across the entire window.
This update step typically occurs when a tracked feature goes out of view of the camera, but it may also be triggered if the length of a feature track exceeds some preset threshold.

@@

\subsection{MSCKF State Correction Equations}
Now that we have estimated the positions of any features to be used in the state update, we can apply the corresponding motion constraints to the window of poses from which each feature was observed.
We begin by forming the exteroceptive measurement error corresponding to an observation $\Vec{z}_i^{(j)}$ of feature $f_j$ from camera pose $C_i$:
\begin{align} \label{eq:residual_def}
    \Vec{r}_i^{(j)} &:= \Vec{z}_i^{(j)} - \hat{\Vec{z}}_i^{(j)}
\end{align}
where
\begin{align}
    \hat{\Vec{z}}_i^{(j)} &= \frac{1}{\hat{Z}^{(j)}_{C_i}} \bbm \hat{X}^{(j)}_{C_i} & \hat{Y}^{(j)}_{C_i} \ebm ^T
\end{align}
with 
\begin{align}
    \hat{\Vec{p}}_{C_i}^{f_j C_i} &= \bbm \hat{X}^{(j)}_{C_i} & \hat{Y}^{(j)}_{C_i} & \hat{Z}^{(j)}_{C_i} \ebm ^T \notag \\
                                 &= \hat{\Vec{C}}_{C_i G} \left( \hat{\Vec{p}}_G^{f_j G} - \hat{\Vec{p}}_G^{C_i G} \right).
\end{align}

If we linearize (\ref{eq:residual_def}) about the estimates for the camera pose and feature position, we obtain an estimate of the exteroceptive measurement error
\begin{align}
    \Vec{r}_i^{(j)} &\simeq \Vec{H}_{\Vec{x}, i}^{(j)} \widetilde{\Vec{x}}_i + \Vec{H}_{f, i}^{(j)} \widetilde{\Vec{p}}_G^{f_j G} + \Vec{n}_i^{(j)}
\end{align}
where $\Vec{H}_{\Vec{x}, i}^{(j)}$ and $\Vec{H}_{f, i}^{(j)}$ are the Jacobians of the measurement of feature $f_j$ from camera pose $C_i$ with respect to the filter state and the position of the feature, respectively.
These are given by
\begin{align}
    \Vec{H}_{\Vec{x}, i}^{(j)} &= \bbm \Vec{0} & \Vec{J}^{\left(j\right)}_i \left(\hat{\Vec{p}}_{C_i}^{f_j C_i}\right)^\times & -\Vec{J}^{\left(j\right)}_i \hat{\Vec{C}}_{C_i G} & \Vec{0} \ebm \\
    \Vec{H}_{f, i}^{(j)} &= \Vec{J}^{\left(j\right)}_i \hat{\Vec{C}}_{C_i G}
\end{align}
where the left $\Vec{0}$ in $\Vec{H}_{\Vec{x}, i}^{(j)}$ has dimension $2\times\left(12+6\left(i-1\right)\right)$, the right $\Vec{0}$ has dimension $2\times6\left(N-i\right)$, and
\begin{align}
    \Vec{J}^{\left(j\right)}_i &= \frac{1}{\left(\hat{Z}^{(j)}_{C_i}\right)^2}
    \bbm \hat{Z}^{(j)}_{C_i} & 0 & -\hat{X}^{(j)}_{C_i} \vspace{0.5em} \\
             0 & \hat{Z}^{(j)}_{C_i} & -\hat{Y}^{(j)}_{C_i}
    \ebm .
\end{align}
$\Vec{n}_i^{(j)}$ is a zero-mean Gaussian noise term with covariance matrix $\Vec{R}_i^{(j)} = \text{diag}\left\{\sigma^2_{u'}, \sigma^2_{v'}\right\}$.

@@

However, the EKF assumes that measurement errors are linear in the error in the state and have an additive zero-mean Gaussian noise component that is \textit{uncorrelated} to the state.
Since the term $\Vec{H}_{f}^{(j)} \widetilde{\Vec{p}}_G^{f_j G}$ \textit{is} correlated to the state, $\Vec{r}^{(j)}$ is not of the correct form for the EKF and must be modified to de-correlate it from the state.
Figure \ref{fig:nullspace} shows the effect of attempting to use $\Vec{r}^{(j)}$ directly in the EKF.
The correlation between the measurement noise and the state causes the filter estimates to drift further from the true values than with a modified, de-correlated measurement error.
As an aside, note that the SWF does not have this problem because feature positions are part of the state itself and are jointly optimized along with the poses.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiments} \label{sec:experiments}
We compared the performance of the MSCKF and SWF algorithms on two intervals in the ``Starry Night'' dataset from the University of Toronto Institute for Aerospace Studies (UTIAS), as well as several runs from the KITTI Dataset \cite{Geiger:2013:IJRR}.
Both datasets include data streams from a stereo camera and IMU.
Since our algorithms are designed to make use of a monocular camera, we artificially blinded the stereo camera by using measurements from the left camera only.

@@

\subsection{Starry Night Dataset}
The ``Starry Night'' dataset consists of a rigidly attached stereo camera and IMU (Figure \ref{fig:sensorhead}) observing a set of 20 features while moving along an arbitrary 3D path.
This dataset is well-suited to evaluating SLAM algorithms since accurate ground truth from a Vicon motion capture system is available for both the sensor head motion and the feature positions.

@@

Figure \ref{fig:comp_1} shows absolute translational and rotational errors for pure IMU integration and various parameter settings for the MSCKF and SWF.
For the MSCKF, we varied both the maximum number of observations before triggering an update and the minimum number of observations for a feature track to be used in an update.
For the SWF, we varied the number of states in the sliding window.
These parameters did not significantly affect the accuracy of either filter, however Figure \ref{fig:comp_1} shows some small gains in accuracy for larger feature track lengths and window sizes.

On this interval, the SWF slightly outperforms the MSCKF in terms of translational accuracy in the $y$- and $z$-directions, and significantly outperforms the MSCKF in terms of rotational accuracy about the $x$- and $y$-axes.
The MSCKF does not perform much better than pure IMU integration on this interval, likely due to the overall low number of feature tracks in this dataset.

% \setcounter{figure}{5}
\begin{figure*}
    \centering
    \subfigure[At least three features are visible in $64\%$ of the interval $k\in(500,1000)$. The number of visible features is frequently exceeds ten.] {
        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-WindowSize-500-1000}
        \label{fig:comp_1}
    }
    ~
    \subfigure[At least three features are visible in only $56\%$ of the interval $k\in(1215,1715)$. The number of visible features rarely exceeds ten.] {
        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-WindowSize-1215-1715}
        \label{fig:comp_2}
    }
    \caption{Comparison of MSCKF, SWF, and IMU integration for multiple parameter settings on two intervals of the ``Starry Night'' dataset. The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature track lengths before an EKF update is triggered, while the numbers next to ``SWF'' refers to the number of states in the sliding window.}
\end{figure*}

@@

Figure \ref{fig:comp_2} shows absolute translational and rotational errors for pure IMU integration and the same MSCKF and SWF parameter settings as in Experiment 1.
Compared to Interval 1, the difference in performance between the MSCKF and SWF is less pronounced, with the exception of the SWF's clearly superior translational accuracy in the $y$-direction.
Neither filter performs substantially better than pure IMU integration on this interval, again likely due to the overall low number of feature tracks in this dataset.

@@

\subsubsection{Effect of feature density}
In order to investigate the effect of feature density on the performance of the MSCKF and SWF, we modified the ``Starry Night'' dataset by creating synthetic feature distributions with larger spatial extents and more features than the original dataset.
We constructed each dataset so that the larger maps contained the same features as the smaller maps, plus additional features to make up the difference.
In each dataset, the IMU data remained unchanged from the original dataset and we corrupted the synthetic camera measurements with zero-mean Gaussian noise.
% Figure \ref{fig:trajectory_groundtruth_100lessnoisy} shows the feature distribution for the 100-feature synthetic dataset.
By generating longer feature tracks and increasing the number of features visible at each time step, these synthetic datasets allowed for a clearer comparison between the MSCKF and SWF.

@@

Figure \ref{fig:comp_3} compares translational and rotational Root Mean Squared Error (RMSE) of the MSCKF, SWF, and pure IMU integration on three synthetic datasets with 40, 60, and 100 features.
With more features, both algorithms consistently outperform pure IMU integration, and the SWF outperforms the MSCKF by a wide margin on all three datasets.

Note that the MSCKF's performance improves as the number of features increases, while the SWF's performance is not significantly affected by increasing feature count.
This result indicates that the MSCKF is much more sensitive to feature density than the SWF.
This may be due to the fact that the MSCKF updates its state whenever a feature goes out of view and does not associate tracks corresponding to the same feature.
If observations of a given feature are frequently interrupted, each track will not constrain the full set of poses from which the feature is visible.
As feature density increases, the number of long feature tracks is likely to increase, and so the MSCKF benefits from more high-quality motion constraints.
Since the SWF always associates all observations of each feature in a given window, it is not sensitive to the contiguity of the feature observations.
% \setcounter{figure}{7}
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Noisy}

    \caption{Comparison of Root Mean Squared Error (RMSE) of the MSCKF, SWF, and pure IMU integration for three different synthetic maps generated based on the IMU data in the dataset interval $k\in[1215,1715]$. The numbers next to ``MSCKF'' and ``SWF'' refer to the number of features in the dataset (40, 60 or 100).}
    \label{fig:comp_3}
\end{figure}

Table \ref{tab:rmse_nees} provides summary statistics of each algorithm's performance on the interval shown in Figure \ref{fig:comp_3}.
As expected, the SWF outperforms the MSCKF in terms of average RMSE on both rotation estimation and position estimation.
It is worth noting, however, that, on average, the MSCKF achieves low Normalized Estimation Error Squared (NEES) values, which indicates that the filter scores well on consistency if not on accuracy.
The average NEES values for the SWF are substantially higher because the SWF treats each window independently and has no built-in mechanism for propagating uncertainty from window to window (\cite{Sibley:2010:JFR} describes how this can be remedied by marginalizing old poses, a technique which was not implemented in this work).
The accumulated drift error in the SWF quickly escapes the estimator's reported uncertainty envelope, and so this form of the SWF may be considered inconsistent in the absence of a known map.

@@

\begin{table}
    \centering
    \caption{Comparison of Average Root Mean Squared Error (ARMSE), Average Normalized Estimation Error Squared (ANEES), and Compute Time of IMU integration, MSCKF, and SWF on synthetic datasets for the interval $k\in[1215,1715]$.}
    \begin{threeparttable}
    \begin{tabular}{llccc} 
                    &                   & \multicolumn{3}{c}{Feature Count} \B      \\ \cline{3-5}
                    &                   & 40        & 60        & 100       \T\B    \\ \hline
        IMU Only    & Position ARMSE    & 0.3679    & 0.3679    & 0.3679    \T      \\
                    & Rotation ARMSE    & 0.1452    & 0.1452    & 0.1452            \\
                    & ANEES             & 0.2850    & 0.2850    & 0.2850            \\
                    & Compute Time\tnote{\dag}      & 8.90 s    & 8.90 s    & 8.90 s    \B      \\ \hline
        MSCKF (20-100)       & Position ARMSE    & 0.2672    & 0.2550    & 0.2304    \T      \\
                    & Rotation ARMSE    & 0.1378    & 0.1247    & 0.0952            \\
                    & ANEES             & 10.18     & 12.03     & 16.76             \\
                    & Compute Time\tnote{\dag}     & 12.19 s   & 14.64 s   & 20.58 s   \B      \\ \hline
        SWF (25)         & Position ARMSE    & 0.1750    & 0.1687    & 0.1755    \T      \\
                    & Rotation ARMSE    & 0.0495    & 0.0377    & 0.0481            \\
                    & ANEES             & 2280      & 2093      & 2013              \\
                    & Compute Time\tnote{\dag}     & 114.3 s   & 175.9 s   & 245.3 s   \B      \\ \hline
    \end{tabular}
    \begin{tablenotes}
        \item [\dag] Running MATLAB 2014b on a MacBook Pro Retina (11,3) with a 2.3GHz Intel Core i7 processor and 16 GB of DDR3L RAM.
    \end{tablenotes}
    \end{threeparttable}
    \label{tab:rmse_nees}
    % \vspace{-0.4cm}
\end{table}

\subsection{KITTI Dataset}
\begin{figure*}
    \centering
    \subfigure[RMSE on the \texttt{2011\_09\_26\_drive\_0001} traverse.] {
        \centering
        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_2011_09_26_drive_0001}
    }
    ~
    \subfigure[RMSE on the \texttt{2011\_09\_26\_drive\_0036} traverse.] {
        \centering
        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_2011_09_26_drive_0036}
    }

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Accurate and consistent egomotion estimation is an important part of autonomous navigation.
For this task, the combination of visual and inertial sensors is an inexpensive, compact, and complementary hardware suite that can be used on many ground and aerial vehicles.
In this work, we compare two modern approaches to egomotion estimation: the Multi-State Constraint Kalman Filter (MSCKF) and the Sliding Window Filter (SWF).
Both filters use an Inertial Measurement Unit (IMU) to estimate the motion of a vehicle and then correct this estimate with observations of salient features from a monocular camera.
While the SWF estimates feature positions as part of the filter state itself, the MSCKF optimizes feature positions in a separate procedure without including them in the filter state.
We present experimental characterizations and comparisons of the MSCKF and SWF on data from a moving hand-held sensor rig, as well as several traverses from the KITTI dataset.
In particular, we compare the accuracy and consistency of the two filters, and analyze the effect of feature track length and feature density on the performance of each filter.
In general, our results show the SWF to be more accurate than the MSCKF.
However, the MSCKF is significantly more computationally efficient, has good consistency properties, and improves in accuracy as more features are tracked. 
\end{abstract}

@@

The question, then, is how best to fuse measurements from these two sensor types to arrive at an accurate estimate of a vehicle's motion over time.
This problem is often complicated by the absence of a known map of features from which the camera can generate measurements.
Any solution must therefore solve a Simultaneous Localization and Mapping (SLAM) problem, although the importance placed on the mapping component may vary from algorithm to algorithm.
In this work we characterize, compare, and contrast the performance of two modern solutions to the visual-inertial SLAM problem, the Sliding Window Filter (SWF) and the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}, on data from a moving hand-held sensor rig, as well as several traverses from the KITTI dataset \cite{Geiger:2013:IJRR}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sliding Window Filter} \label{sec:SWF}
In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and feature positions.
The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton (GN) optimization over a sliding window of $K$ poses $\Vec x_{k\in[1,K]}$ observing $M$ features $f_{j\in[1,M]}$: 
\begin{equation}
\Vec{x} := \bbm \Vec x_0^T & \hdots & \Vec x_K^T & {\Vec {p}_G^{f_1G \hspace{0.3em} T}} & \hdots & {\Vec {p}_G^{f_MG \hspace{0.3em} T}} \ebm ^T.
\end{equation}
In our notation, $\Vec p^{BA}_A$ is the vector from the origin of frame $\mathcal{F}_A$ to the origin of frame $\mathcal{F}_B$ expressed in $\mathcal{F}_A$, and $\Vec C_{BA}$ is the rotation matrix from $\mathcal{F}_A$ to $\mathcal{F}_B$.

@@

An important advantage of the SWF is that its computational cost depends on the number of features in the current window rather than the number of features in the entire map.
By varying the extent of the sliding window, the computational cost of the algorithm can be tailored to fit a given compute envelope, which makes the algorithm suitable for online operation over paths of arbitrary length.

However, the hard cut-off of the SWF may result in only a subset of measurements of each feature contributing to the optimization.
As a result, the filter may not maximally constrain some poses, and hence localization may be less accurate than we could expect from the full batch solution.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Multi-State Constraint Kalman Filter} \label{sec:MSCKF}
The Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
The key idea of the MSCKF is to maintain a variable window of vehicle poses and to simultaneously update each pose in the window using batch-optimized estimates of features that are visible across the entire window.
This update step typically occurs when a tracked feature goes out of view of the camera, but it may also be triggered if the length of a feature track exceeds some preset threshold.

@@

\subsection{MSCKF State Correction}
Now that we have estimated the positions of any features to be used in the state update, we can apply the corresponding motion constraints to the window of poses from which each feature was observed.
We begin by forming the exteroceptive measurement error corresponding to an observation $\Vec{z}_i^{(j)}$ of feature $f_j$ from camera pose $C_i$:
\begin{align} \label{eq:residual_def}
    \Vec{r}_i^{(j)} &:= \Vec{z}_i^{(j)} - \hat{\Vec{z}}_i^{(j)}
\end{align}
where
\begin{align}
    \hat{\Vec{z}}_i^{(j)} &= \frac{1}{\hat{Z}^{(j)}_{C_i}} \bbm \hat{X}^{(j)}_{C_i} & \hat{Y}^{(j)}_{C_i} \ebm ^T
\end{align}
with 
\begin{align}
    \hat{\Vec{p}}_{C_i}^{f_j C_i} &= \bbm \hat{X}^{(j)}_{C_i} & \hat{Y}^{(j)}_{C_i} & \hat{Z}^{(j)}_{C_i} \ebm ^T \notag \\
                                 &= \hat{\Vec{C}}_{C_i G} \left( \hat{\Vec{p}}_G^{f_j G} - \hat{\Vec{p}}_G^{C_i G} \right).
\end{align}

If we linearize (\ref{eq:residual_def}) about the estimates for the camera pose and feature position, we obtain an estimate of the exteroceptive measurement error
\begin{align}
    \Vec{r}_i^{(j)} &\simeq \Vec{H}_{\Vec{x}, i}^{(j)} \widetilde{\Vec{x}}_i + \Vec{H}_{f, i}^{(j)} \widetilde{\Vec{p}}_G^{f_j G} + \Vec{n}_i^{(j)}
\end{align}
where $\Vec{H}_{\Vec{x}, i}^{(j)}$ and $\Vec{H}_{f, i}^{(j)}$ are the Jacobians of the measurement of feature $f_j$ from camera pose $C_i$ with respect to the filter state and the position of the feature, respectively:
\begin{align}
    \Vec{H}_{\Vec{x}, i}^{(j)} &= \bbm \Vec{0} & \Vec{J}^{\left(j\right)}_i \left(\hat{\Vec{p}}_{C_i}^{f_j C_i}\right)^\times & -\Vec{J}^{\left(j\right)}_i \hat{\Vec{C}}_{C_i G} & \Vec{0} \ebm \\
    \Vec{H}_{f, i}^{(j)} &= \Vec{J}^{\left(j\right)}_i \hat{\Vec{C}}_{C_i G}
\end{align}
where the left $\Vec{0}$ in $\Vec{H}_{\Vec{x}, i}^{(j)}$ has dimension $2\times\left(12+6\left(i-1\right)\right)$, the right $\Vec{0}$ has dimension $2\times6\left(N-i\right)$, and
\begin{align}
    \Vec{J}^{\left(j\right)}_i &= \frac{1}{\left(\hat{Z}^{(j)}_{C_i}\right)^2}
    \bbm \hat{Z}^{(j)}_{C_i} & 0 & -\hat{X}^{(j)}_{C_i} \vspace{0.5em} \\
             0 & \hat{Z}^{(j)}_{C_i} & -\hat{Y}^{(j)}_{C_i}
    \ebm .
\end{align}
$\Vec{n}_i^{(j)}$ is a zero-mean Gaussian noise term with covariance matrix $\Vec{R}_i^{(j)} = \text{diag}\left\{\sigma^2_{u'}, \sigma^2_{v'}\right\}$.

@@

However, the EKF assumes that measurement errors are linear in the error in the state and have an additive zero-mean Gaussian noise component that is \textit{uncorrelated} to the state.
Since the term $\Vec{H}_{f}^{(j)} \widetilde{\Vec{p}}_G^{f_j G}$ \textit{is} correlated to the state, $\Vec{r}^{(j)}$ is not of the correct form for the EKF and must be modified to de-correlate it from the state.
Figure \ref{fig:nullspace} shows the effect of attempting to use $\Vec{r}^{(j)}$ directly in the EKF.
The correlation between the measurement noise and the state causes the filter estimates to drift further from the true values than with a modified, de-correlated measurement error.
As an aside, we note that the SWF does not have this problem because feature positions are jointly optimized along with the poses.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiments} \label{sec:experiments}
We compared the performance of the MSCKF and SWF on two intervals in the ``Starry Night'' dataset from the University of Toronto Institute for Aerospace Studies (UTIAS), as well as 530 m of urban driving from the KITTI Dataset \cite{Geiger:2013:IJRR}.
Both datasets include data streams from a stereo camera and IMU.
Since our algorithms are designed to make use of a monocular camera, we artificially blinded the stereo camera by using images from the left camera only.

@@

\subsection{Starry Night Dataset}
The ``Starry Night'' dataset consists of a rigidly attached stereo camera and IMU (Figure \ref{fig:sensorhead}) observing a set of 20 features while moving along an arbitrary 3D path.
This dataset is well-suited to evaluating SLAM algorithms since ground truth from a Vicon motion capture system is available for both the sensor head motion and the feature positions.
We conducted three experiments on this dataset to compare the effect of feature visibility and window size on each algorithm.
We discuss each of these in turn.

\begin{figure*}
    \centering
    \subfigure[At least three features are visible in $64\%$ of the interval $(500,1000)$. The number of visible features is frequently exceeds ten.] {
        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-WindowSize-500-1000}
        \label{fig:comp_1}
    }
    ~
    \subfigure[At least three features are visible in only $56\%$ of the interval $(1215,1715)$. The number of visible features rarely exceeds ten.] {
        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-WindowSize-1215-1715}
        \label{fig:comp_2}
    }
    \caption{Comparison of MSCKF, SWF, and IMU integration for multiple parameter settings on two intervals of the ``Starry Night'' dataset. The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature track lengths before an EKF update is triggered, while the numbers next to ``SWF'' refers to the number of states in the sliding window.}
\end{figure*}

@@

Figure \ref{fig:comp_1} shows translational and rotational root mean squared errors (RMSE) for pure IMU integration and various parameter settings for the MSCKF and SWF.
For the MSCKF, we varied the maximum number of observations before triggering an update and the minimum number of observations for a feature track to be used in an update.
For the SWF, we varied the number of states in the window.
These parameters did not significantly affect the accuracy of either filter, however Figure \ref{fig:comp_1} shows some small gains in accuracy for larger feature track lengths and window sizes.

On this interval, the SWF outperforms the MSCKF in terms of both translational and rotational RMSE.
The MSCKF does not perform much better than pure IMU integration on this interval, likely due to the overall low number of feature tracks in this dataset.

@@

Figure \ref{fig:comp_2} shows translational and rotational RMSE for pure IMU integration and the same MSCKF and SWF parameter settings as in Experiment 1.
Compared to Interval 1, the difference in performance between the MSCKF and SWF is less pronounced, with the exception of the SWF's clearly superior translational accuracy in the $y$-direction.
Neither filter performs substantially better than pure IMU integration on this interval, again likely due to the overall low number of feature tracks in this dataset.

@@

\subsubsection{Effect of feature density}
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Noisy}

    \caption{Comparison of Root Mean Squared Error (RMSE) of the MSCKF, SWF, and pure IMU integration for three synthetic maps generated using the IMU data in the interval $(1215,1715)$. The numbers next to ``MSCKF'' and ``SWF'' refer to the number of features in the dataset (40, 60 or 100).}
    \label{fig:comp_3}
    \vspace{-0.4cm}
\end{figure}

In order to investigate the effect of feature density on the performance of the MSCKF and SWF, we modified the ``Starry Night'' dataset by creating synthetic feature distributions with larger spatial extents and more features than the original dataset.
We constructed each dataset so that the larger maps contained the same features as the smaller maps, plus additional features to make up the difference.
In each dataset, we retained the IMU data from the original dataset and corrupted the synthetic camera measurements with zero-mean Gaussian noise.
% Figure \ref{fig:trajectory_groundtruth_100lessnoisy} shows the feature distribution for the 100-feature synthetic dataset.
By generating longer feature tracks and increasing the number of features visible at each time step, these synthetic datasets allowed for a clearer comparison between the MSCKF and SWF.

@@

Figure \ref{fig:comp_3} compares translational and rotational RMSE of the MSCKF, SWF, and pure IMU integration on three synthetic datasets with 40, 60, and 100 features.
With more features, both algorithms consistently outperform pure IMU integration, and the SWF outperforms the MSCKF by a wide margin on all three datasets.

Note that the MSCKF's performance improves as the number of features increases, while the SWF's performance is not significantly affected by increasing feature count.
This result indicates that the MSCKF is much more sensitive to feature density than the SWF.
This may be due to the fact that the MSCKF updates its state whenever a feature goes out of view and does not associate tracks corresponding to the same feature.
If observations of a given feature are frequently interrupted, each track will not constrain the full set of poses from which the feature is visible.
As feature density increases, the number of long feature tracks is likely to increase, and so the MSCKF benefits from more high-quality motion constraints.
Since the SWF always associates all observations of each feature in a given window, it is not sensitive to the contiguity of the feature observations.

Table \ref{tab:rmse_nees} summarizes each algorithm's performance on the interval shown in Figure \ref{fig:comp_3}.
As expected, the SWF outperforms the MSCKF in terms of average RMSE on both rotation and translation.
It is worth noting, however, that, on average, the MSCKF achieves low Normalized Estimation Error Squared (NEES) values, which indicates that the filter scores well on consistency if not on accuracy.
The average NEES values for the SWF are substantially higher because the SWF treats each window independently and has no built-in mechanism for propagating uncertainty from window to window (\cite{Sibley:2010:JFR} describes how this can be remedied by marginalizing old poses, a technique which was not implemented in this work).
The accumulated drift error in the SWF quickly escapes the estimator's reported uncertainty envelope, and so this form of the SWF may be considered inconsistent in the absence of a known map.

@@

\begin{table}
    \centering
    \caption{Comparison of Average Root Mean Squared Error (ARMSE), Average Normalized Estimation Error Squared (ANEES), and Compute Time of IMU integration, MSCKF, and SWF on synthetic datasets for the interval $(1215,1715)$.}
    \begin{threeparttable}
    \begin{tabular}{llccc} 
                    &                   & \multicolumn{3}{c}{Feature Count} \B      \\ \cline{3-5}
                    &                   & 40        & 60        & 100       \T\B    \\ \hline
        IMU Only    & Trans. ARMSE    & 0.3679    & 0.3679    & 0.3679    \T      \\
                    & Rot. ARMSE    & 0.1452    & 0.1452    & 0.1452            \\
                    & ANEES             & 0.2850    & 0.2850    & 0.2850            \\
                    & Compute Time\tnote{\dag}      & 8.90 s    & 8.90 s    & 8.90 s    \B      \\ \hline
        MSCKF       & Trans. ARMSE    & 0.2672    & 0.2550    & 0.2304    \T      \\
        (20-100)            & Rot. ARMSE    & 0.1378    & 0.1247    & 0.0952            \\
                    & ANEES             & 10.18     & 12.03     & 16.76             \\
                    & Compute Time\tnote{\dag}     & 12.19 s   & 14.64 s   & 20.58 s   \B      \\ \hline
        SWF         & Trans. ARMSE    & 0.1750    & 0.1687    & 0.1755    \T      \\
         (25)       & Rot. ARMSE    & 0.0495    & 0.0377    & 0.0481            \\
                    & ANEES             & 2280      & 2093      & 2013              \\
                    & Compute Time\tnote{\dag}     & 114.3 s   & 175.9 s   & 245.3 s   \B      \\ \hline
    \end{tabular}
    \begin{tablenotes}
        \item [\dag] Running MATLAB 2014b on a MacBook Pro Retina (11,3) with a 2.3GHz Intel Core i7 processor and 16 GB of DDR3L RAM.
    \end{tablenotes}
    \end{threeparttable}
    \label{tab:rmse_nees}
    % \vspace{-0.4cm}
\end{table}

\subsection{KITTI Dataset}
In addition to the ``Starry Night'' datasets, we also compared the performance of the MSCKF and SWF over several traverses from the KITTI dataset \cite{Geiger:2013:IJRR} totalling $530$ m of urban driving.
To accommodate our alternative state parametrization, we used the pre-integrated linear velocity measurements provided in the dataset instead of the raw linear accelerations from the IMU.
Figure \ref{fig:kitti} shows the RMSE of pure IMU integration, the MSCKF, and the SWF over each of the four traverses we tested, representative images from which are shown in Figure \ref{fig:composite-kitti}.
Similarly to the ``Starry Night'' results, both algorithms outperformed pure IMU integration, and the SWF slightly outperformed MSCKF in terms of rotational and translational RMSE.
As expected, the MSCKF produced estimates that were more consistent than those of the SWF.

\begin{figure*}
    \centering
    \subfigure[RMSE on KITTI traverse \texttt{0001}.] {
        \centering
        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_2011_09_26_drive_0001}
    }
    ~
    \subfigure[RMSE on KITTI traverse \texttt{0036}.] {
        \centering
        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_2011_09_26_drive_0036}
    }
